---
title: "Tidyverse & Reshaping Data"
date: 2021-03-05
author: "Louise Huuki"
output: html_notebook
---

### What is the Tidyverse?
The "Tidyverse" is a collection of 8 packages designed for data science, and 
designed to work together. https://www.tidyverse.org/

1. ggplot2 - plots data  
2. tibble - "upgraded" data frame  
3. tidyr - data transformations (make data tidy)  
4. readr - import data in tidyverse friendly way  
5. purrr - "upgraded" loops  
6. dplyr - data manipulations (change data)  
7. stringr - work with strings  
8. forcats - work with factors  

```{r}
library(tidyverse)
```

### %>%
* The "pipe" symbol
* from magrittr package - but loaded by tidyverse
* Strings together functions 
  + makes your code more readable
  + fewer intermediate objects
  + helps eliminate parentheses mess
  
https://twitter.com/andrewheiss/status/1359583543509348356
  
```{r}
data <- runif(10)

sqrt(mean(data^2))
data^2 %>% mean() %>% sqrt()
```
  
### dplyr tools
```{r}

diamonds %>% filter(cut  == "Ideal") %>%
  group_by(color) %>%
  summarize(mean_price = mean(price),
            n = n(),
            max_price = max(price)) %>%
  arrange(-mean_price)

```

#### SE example
```{r}
library(DeconvoBuddies)
library(SummarizedExperiment)

rse_bulk_test

pd <- as.data.frame(colData(rse_bulk_test))

head(pd)

table(pd$Sex, pd$Dx)

pd %>% group_by(Sex, Dx) %>%
  summarize(mean_age = mean(Age),
            n = n(),
            min_age = min(Age)
            ) %>%
  # filter(Dx == "Case")  %>%
  identity()



```

## Reshaping 

### Long vs. Wide data

#### Reshape2 vs. tidyr pivot functions

* make longer: melt & pivot_longer
* make wider: dcast & pivot_wider

#### Make Longer
https://twitter.com/bmwiernik/status/1366426648921391105

```{r}
library(reshape2)

est_prop_melt <- est_prop %>%
  rownames_to_column("sample") %>%
  melt(id = "sample")

est_prop_long <- est_prop %>%
  rownames_to_column("sample") %>%
  pivot_longer(!sample, names_to = "cell_type", values_to = "prop")

est_prop_long
```

```{r}
est_prop_long %>% 
  ggplot(aes(cell_type, prop)) +
  geom_boxplot()

```

#### Make Wider
```{r}
est_prop_long %>%
  dcast(sample~cell_type)

col <- "cell_type"

est_prop_long %>%
  pivot_wider(id_cols = sample, names_from = col, values_from = prop)
```


### Joins
combine est_prop with Dx info from pd
```{r}

pd_prop <- pd %>% filter(Sex == "F") %>%
  left_join(est_prop %>% rownames_to_column("sample"), by = c("RNum" = "sample"))

```


#### rehsape with more complicated data

```{r}
(pd_prop_long <- pd_prop %>% pivot_longer(cols = starts_with("cell_"), names_to = "cell_type", values_to = "prop"))
```

```{r}
pd_prop_long %>%
  ggplot(aes(Age, prop, color = cell_type)) +
  geom_point() +
  facet_wrap(~cell_type) +
  NULL
```

